Multi-target detection with application to cryo-electron microscopy

### 2019-02-19 inverse_problems/inverse_problems.tex
Fine 607 session: rearranged a lot of sections 3 and 4

Specifically, for some parameter $\gamma > 0$, the total number of signal occurrences is drawn from a Poisson distribution with expectation $(N-L+1)\gamma$. For each occurrence, a starting location $i$ is drawn uniformly and independently at random from $\{0, \ldots, N-L\}$. Then, one of the signals $x_1, \ldots, x_K$ is selected independently at random from a fixed distribution and added to the observation $y$ with first entry positioned at $i$. 
By definition of the Poisson process, the entries of $s$ are independent and follow a Poisson distribution with expectation $\gamma$. In particular, this model allows for (additive) signal overlap. In Section~\ref{sec:poisson}, we show that the autocorrelations of $y$ are nonetheless equivalent to those arising under our well-separated model above.

@@

Let $x$ be the random variable corresponding to signal occurrences in $y$. In particular, for the model~\eqref{eq:model}, $x=x_k$ with probability $\pi_k$.
Under the separation condition~\eqref{eq:spacing}, the relation between autocorrelations of the observation $y$ and those of $x$ is particularly simple, as we now show. It is useful to introduce some notation: let $\vert s\vert = \sum_i s[i]$ denote the number of signal occurrences in $y$, and let
\begin{equation}
\gamma  = \frac{|s| L}{N}.
\label{eq:gamma}
\end{equation}
In the well-separated model, $\gamma$ is the fraction of entries of $y$ occupied by signal occurrences. Henceforth, we call it the signal density.
The separation condition  imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$.

@@

In our problem setup, we are given one observation (one realization) of $y$. Thus, we cannot compute the autocorrelations of $y$ exactly as they involve taking an expectation against the distribution of $y$. However, by the law of large numbers, as $N$ grows to infinity with $\gamma$ remaining constant, the empirical autocorrelations of $y$ converge to the actual autocorrelations of $y$, that is, % (ergodicity)
\begin{align}
	\lim_{N \to \infty, \gamma \textrm{ constant}} \frac{1}{N} \sum_{i=-\infty}^{\infty} y[i]y[i+\ell_1]\cdots y[i+\ell_{q-1}] = a_y^q[\ell_1, \ldots, \ell_{q-1}].
\end{align}
This provides a concrete means of estimating the quantities $a_y^ q$.

In particular, for $x$ sampled from $\{ x_1, \ldots, x_K \}$ with probabilities $(\pi_1, \ldots, \pi_K)$, the autocorrelations are given in explicit form as:
\begin{equation}
a_x^q  =  \sum_{k=1}^{K} \pi_ka_{x_k}^q.
\end{equation}
We show in Section~\ref{sec:theory_homogeneous} how to estimate $\gamma$ from observations for the particular case where $K = 1$ (all signal occurrences are the same). Furthermore, in Section~\ref{sec:poisson} we show that the autocorrelations under the Poisson model provide the same information about $x$ as those presented in this section.

@@

\subsection{Theory for the homogeneous case} \label{sec:theory_homogeneous}
In this section, we derive some theoretical results for the homogeneous case. \TODO{Note: I am using terms interchangeably}

A signal is determined uniquely by its second- and third-order autocorrelations. Indeed, assuming $z[0]$ and $z[L-1]$ are nonzero (otherwise, redefine the length of the signal), we can recover $z$ explicitly using this identity for $k = 0, \ldots, L-1$:
%
\begin{equation}
%
z[k]  = \frac{z[0]z[k]z[L-1]}{z[0]z[L-1]} = \frac{a_z^3[k,L-1]}{a_z^2[L-1]}.
\label{eq-uniqueness}
%
\end{equation}
This proves the following useful fact:
\begin{proposition} \label{prop:uniqueness}
	%
	A signal $z\in\RL$ is determined uniquely from  $a_z^2$ and $a_z^3$.
\end{proposition}

@@

\begin{proposition} \label{prop:gamma}
	Let $\sigma > 0$ be fixed and assume that the separation condition~\eqref{eq:spacing} holds. If the mean of $x$ is nonzero, then 
	\begin{equation}
	\gamma  \aseq \lim_{N \to \infty}\frac{L (a^1_y)^2}{a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\sigma^2}.
	\end{equation}
\end{proposition}

@@

\subsection{Equivalence between the autocorrelations of under the separation and the Poisson process \TODO{?}} \label{sec:poisson}

In this section we omit the affect of the noise as it remains the same for the well-separated model. We aim to show that the moments of

\begin{corollary} \TODO{re-write}
	From the first three moments of the Poisson process model, one can recover the first three moments from the strongly-separated model.
\end{corollary}

This uniqueness result for multi-dimensional signals (especially for 2-D images) is the basis of a popular imaging technique called coherent diffraction imaging (CDI). In CDI, an object is
illuminated with a coherent wave, and the far field diffraction
intensity pattern  is measured, corresponding object's Fourier magnitude~\cite{miao1999extending,shechtman2015phase}. 
If the support of the object is known and less than half of the support of the measured signal, then the data is equivalent to the second-order autocorrelation. The computational problem of recovering the signal is usually referred to as \emph{phase retrieval} or \emph{phase problem}.
Albeit the uniqueness result, recently it has been shown that, at least for 2-D images, the problem is  ill-conditioned~\cite{barnett2018geometry}. That is, there exist different images whose second-order autocorrelation agree up to machine precision.

\subsection{How many signals can be recovered in the heterogeneous model?}
\label{sec:heterogeneity}

@@

>>>>

\TODO{Careful that $\gamma$ changed by a factor of $L$: need to modify here too.} Specifically, for some parameter $\gamma > 0$, the total number of signal occurrences is drawn from a Poisson distribution with expectation $(N-L+1)\gamma$. For each occurrence, a starting location $i$ is drawn uniformly and independently at random from $\{0, \ldots, N-L\}$. Then, one of the signals $x_1, \ldots, x_K$ is selected independently at random from a fixed distribution and added to the observation $y$ with first entry positioned at $i$. 
By definition of the Poisson process, the entries of $s$ are independent and follow a Poisson distribution with expectation $\gamma$. In particular, this model allows for (additive) signal overlap. In Section~\ref{sec:poisson}, we show that the autocorrelations of $y$ are nonetheless equivalent to those arising under our well-separated model above.

@@

In what follows, $x$ is the random variable corresponding to signal occurrences in $y$. In particular, for the model~\eqref{eq:model}, 
%$x=x_k$ with probability $\pi_k$.
%In particular, for
$x$ is sampled from $\{ x_1, \ldots, x_K \}$ with probabilities $(\pi_1, \ldots, \pi_K)$, so that its autocorrelations are given in explicit form as:
\begin{equation}
	a_x^q  = \sum_{k=1}^{K} \pi_ka_{x_k}^q.
\end{equation}
We are given one observation (one realization) of $y$. Thus, we cannot compute the autocorrelations of $y$ exactly as they involve taking an expectation against the distribution of $y$. However, by the law of large numbers, as $N$ grows to infinity with $\gamma$ remaining constant, the empirical autocorrelations of $y$ converge to the actual autocorrelations of $y$, that is, % (ergodicity)
\begin{align}
\lim_{N \to \infty, \gamma \textrm{ constant}} \frac{1}{N} \sum_{i=-\infty}^{\infty} y[i]y[i+\ell_1]\cdots y[i+\ell_{q-1}] = a_y^q[\ell_1, \ldots, \ell_{q-1}].
\end{align}
This provides a concrete means of estimating the quantities $a_y^ q$. In the remainder of this section, we relate the observables $a_y^q$  to the unknowns $a_x^q$, first under the well-separated model, then under the Poisson model.

\subsection{Autocorrelations under the well-separated model}

Under the separation condition~\eqref{eq:spacing}, the relation between autocorrelations of the observation $y$ and those of $x$ is particularly simple, as we now show. It is useful to introduce some notation: let $\vert s\vert = \sum_i s[i]$ denote the number of signal occurrences in $y$, and let
\begin{equation}
\gamma  = \frac{|s| L}{N}.
\label{eq:gamma}
\end{equation}
This $\gamma$ is the fraction of entries of $y$ occupied by signal occurrences. Henceforth, we call it the signal density.
The separation condition  imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$.

@@

\TODO{We show in Section~\ref{sec:theory_homogeneous} how to estimate $\gamma$ from observations for the particular case where $K = 1$ (all signal occurrences are the same).}

%Furthermore, in Section~\ref{sec:poisson} we show that the autocorrelations under the Poisson model provide the same information about $x$ as those presented in this section.

\subsection{Autocorrelations under the Poisson model}

\begin{comment}
In this section we omit the affect of the noise as it remains the same for the well-separated model. We aim to show that the moments of

\begin{corollary} \TODO{re-write}
	From the first three moments of the Poisson process model, one can recover the first three moments from the strongly-separated model.
\end{corollary}
\end{comment}

\TODO{What we want to have in this section}
\begin{enumerate}
	\item Recall the Poisson model (it's already described at the end of intro, so keep it short).
	\item Define a proper notion of $\gamma$ (careful with scaling $L$).
	\item State formulas for $a_y^q$ with $q = 1, 2, 3$ (ideally, including noise terms) to parallel what we did in the well-separated section. It would be better to do everything with autocorrelations rather than moments to keep the story uniform.
	\item (optional) Claim: under some assumptions (knowing $\gamma$? Others?), the autocorrelations of $y$ of order $q = 1, 2, 3$ under the Poisson model provide the same information about $x$ as the corresponding autocorrelations under the well-separated model. Could also be in Section 4.1.
\end{enumerate}

@@

\TODO{Explain difference between homogeneous and heterogeneous}

\subsection{Guarantees for the homogeneous case} \label{sec:theory_homogeneous}
In this section, we derive some theoretical results for the homogeneous case. \TODO{Note: I am using terms interchangeably}

A signal is determined uniquely by its second- and third-order autocorrelations. Indeed, assuming $z[0]$ and $z[L-1]$ are nonzero (otherwise, redefine the length of the signal), we can recover $z$ explicitly using this identity for $k = 0, \ldots, L-1$:
%
\begin{equation}
%
z[k]  = \frac{z[0]z[k]z[L-1]}{z[0]z[L-1]} = \frac{a_z^3[k,L-1]}{a_z^2[L-1]}.
\label{eq-uniqueness}
%
\end{equation}
This proves the following useful fact:
\begin{proposition} \label{prop:uniqueness}
	%
	A signal $z\in\RL$ is determined uniquely from  $a_z^2$ and $a_z^3$. \TODO{Definitely only clear for the homogeneous case}
\end{proposition}

@@

\begin{proposition} \label{prop:gamma}
	\TODO{Can we do this simultaneously for Poisson? or do we only need to do it for well-separated and invoke the equivalence claim? Not if that claim requires knowledge of $\gamma, \sigma$.}
	Let $\sigma > 0$ be fixed and assume that the separation condition~\eqref{eq:spacing} holds. If the mean of $x$ is nonzero, then 
	\begin{equation}
	\gamma  \aseq \lim_{N \to \infty}\frac{L (a^1_y)^2}{a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\sigma^2}.
	\end{equation}
\end{proposition}

@@

\subsection{Elemetary limitations of the heterogeneous case}
\label{sec:heterogeneity}

\TODO{Check and specify if this holds for both models}

@@

\TODO{Debate this later}

This uniqueness result for multi-dimensional signals (especially for 2-D images) is the basis of a popular imaging technique called coherent diffraction imaging (CDI). In CDI, an object is
illuminated with a coherent wave, and the far field diffraction
intensity pattern  is measured, corresponding object's Fourier magnitude~\cite{miao1999extending,shechtman2015phase}. 
If the support of the object is known and less than half of the support of the measured signal, then the data is equivalent to the second-order autocorrelation. The computational problem of recovering the signal is usually referred to as \emph{phase retrieval} or \emph{phase problem}.
Despite the uniqueness result, recently it has been shown that, at least for 2-D images, the problem is  ill-conditioned~\cite{barnett2018geometry}. That is, there exist different images whose second-order autocorrelation agree up to machine precision.
